---
title: "Quick start guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


The objective of this package is to compute rates adjusted by a reference population or other rate. This is a very common procedure in epidemiology, allowing the comparison of rates of a event (like mortality) among groups that have different age distributions. 

Some packages like the `epitools` compute these adjusted rates. This package functions wraps the `epitools` functions in a tidy way, allowing the computation of age adjusted rates for several groups using key variables, like year and regions for example.

## Setup

### Installing the package

```{r eval=FALSE}
devtools::install_github("rfsaldanha/tidyrates")
```

### Load the package

```{r}
library(tidyrates)
```

## Data

### Events and population data

Let's use the Fleiss dataset, quoted by the `epitools ` package (Fleiss, 1981, p. 249 ). 

```{r}
population <- c(230061, 329449, 114920, 39487, 14208, 3052,
72202, 326701, 208667, 83228, 28466, 5375, 15050, 175702,
207081, 117300, 45026, 8660, 2293, 68800, 132424, 98301, 
46075, 9834, 327, 30666, 123419, 149919, 104088, 34392, 
319933, 931318, 786511, 488235, 237863, 61313)

population <- matrix(population, 6, 6, 
dimnames = list(c("Under 20", "20-24", "25-29", "30-34", "35-39",
"40 and over"), c("1", "2", "3", "4", "5+", "Total")))

count <- c(107, 141, 60, 40, 39, 25, 25, 150, 110, 84, 82, 39,
3, 71, 114, 103, 108, 75, 1, 26, 64, 89, 137, 96, 0, 8, 63, 112,
262, 295, 136, 396, 411, 428, 628, 530)
count <- matrix(count, 6, 6, 
dimnames = list(c("Under 20", "20-24", "25-29", "30-34", "35-39",
"40 and over"), c("1", "2", "3", "4", "5+", "Total")))
```

```{r}
population
```

```{r}
count
```

The Fleiss data present events (`count` object) and population (`population` object) for six age groups on five different groups (from 1 to 5+).

The `tidyrates` package present the same Fleiss data in a tidy way, with a tibble in long format.

```{r}
fleiss_data
```

The `key` variable refers to the groups, `age_group` to the age groups, `name` separates the `values` into events and population. 

*You may use this same structure for your use case data.*

### Reference population data

The Fleiss example uses the average population as standard population reference. 

```{r}
standard<-apply(population[,-6], 1, mean)
standard
```

Using `tidyrates`, we must supply a tibble with two variables: age group and population.

```{r}
standard_pop <- tibble::tibble(
    age_group = c("Under 20", "20-24", "25-29", "30-34", "35-39", "40 and over"),
    pop = c(63986.6, 186263.6, 157302.2, 97647.0, 47572.6, 12262.6)
  )
```


## Direct adjusted rate

To use the direct adjustment procedure, `tidyrate` present the `rate_adj_direct` function. The `.data` argument must be a tibble with the events and population data, and the `.std` argument must be standard population tibble. The `.keys` argument must point to the groups variables on the `.data` tibble.

The `rate_adj_direct` will compute the crude rate, adjusted rate and exact confidence intervals for each group.

```{r}
rate_adj_direct(fleiss_data, .std = standard_pop, .keys = "key")
```

